### 2017-05-30 doc/categories.tex

Figure~\ref{fig:top_words} shows the per-dependency 
omission score distributions for seven top-ranked words.
There are clear and large differences in how these words
impact the network's representation depending on what grammatical
function they fulfill. They all have large omission scores when they
occur as {\sc nsubj} (nominal subject) or {\sc root}, likely due to the fact that these
grammatical functions typically have a large contribution to the
complete meaning of a sentence.  Conversely, all have small omission 
scores when appearing as {\sc conj} (conjunct): this is probably because in this position
they share their contribution with the first, often more important,
member of the conjunction, for example: {\it A cow and its baby eating
  grass}. 
% The pattern for {\sc nn} (nominal modifier) is a bit more complicated: for
% four of the words shown (as well as for most other words not shown in
% the figure), the score is very low in this grammatical function--presumably 
% because most words contribute less to the sentence meaning when used  
% as modifiers than as heads (e.g.\ {\it a clock tower}). However,
% for the words {\it zebra} and {\it water}, omission scores are high
% when they act as a nominal modifier {\sc nn}. This appears due to two reasons:
% \begin{enumerate}
% \item For {\it zebra}, there are frequent erroneous parses such as {\it
%     zebra/{\sc nn} browsing/{\sc root}} instead of {\it zebra/{\sc nsubj} browsing/{\sc
%     root}}. The network does not make this mistake, and treats these
%     occurrences of {\it zebra} according to its importance as {\sc nsubj}.
% \item {\it Water} as a modifier often changes the meaning of its head in
%  a visually salient way: e.g.\ {\it water fall, water balloon, water scene, water skiing}, and
%   thus the network learns that this particular word is important in the modifier position.
% \end{enumerate}

@@

\begin{figure}
\centering
 \includegraphics[scale=0.4]{position-coef.pdf}
 \caption{Coefficients on the y-axis of {\sc LM full} corresponding to the
position variables on the x-axis.}
 \label{fig:posrqs}
\end{figure}

@@

For the {\sc Visual} model it is less clear what to
expect: on the one hand due to their chain structure,  
RNNs are better at keeping track of
short-distance rather than long-distance dependencies and thus we can expect 
tokens in positions closer to the end of the sentence to be more important. 
On the other hand in English the information structure a single
sentence is 
expressed via linear ordering: the {\sc topic} of a 
sentence appears sentence-initially, and the {\sc comment} follows. 
In the context of other text types such as dialog or multi-sentence \label{edit:topiccomment}
narrative structure, we would expect {\sc comment} to often be more
important than {\sc topic} as {\sc comment} will often 
contain new information in these cases. In our setting of image captions 
however, sentences are not part of larger discourse, it is sentence
initial material that typically contains the most important 
objects depicted in the image, e.g. {\it {\underline{two zebras}} are grazing in tall grass on a savannah.}
Thus, for the task of predicting features of the visual scene, it would 
be advantageous to detect the topic of the sentence and up-weight its
importance in the final meaning representation. Figure \ref{fig:posrqs}
appears to support this hypothesis and the network does 
learn to pay more attention to  words appearing
sentence-initially. This effect seems to be to some extent mixed with the recency
bias of RNNs as perhaps indicated by the relatively high coefficient of the {\it last}
position for {\sc Visual}.

>>>>

Figure~\ref{fig:top_words} shows the per-dependency 
omission score distributions for seven top-ranked words.
There are clear and large differences in how these words
impact the network's representation depending on what grammatical
function they fulfill. They all have large omission scores when they
occur as {\sc nsubj} (nominal subject) or {\sc root}, likely due to the fact that these
grammatical functions typically have a large contribution to the
complete meaning of a sentence.  Conversely, all have small omission 
scores when appearing as {\sc conj} (conjunct): this is probably because in this position
they share their contribution with the first, often more important,
member of the conjunction, for example in {\it A cow and its baby eating
  grass}. 
% The pattern for {\sc nn} (nominal modifier) is a bit more complicated: for
% four of the words shown (as well as for most other words not shown in
% the figure), the score is very low in this grammatical function--presumably 
% because most words contribute less to the sentence meaning when used  
% as modifiers than as heads (e.g.\ {\it a clock tower}). However,
% for the words {\it zebra} and {\it water}, omission scores are high
% when they act as a nominal modifier {\sc nn}. This appears due to two reasons:
% \begin{enumerate}
% \item For {\it zebra}, there are frequent erroneous parses such as {\it
%     zebra/{\sc nn} browsing/{\sc root}} instead of {\it zebra/{\sc nsubj} browsing/{\sc
%     root}}. The network does not make this mistake, and treats these
%     occurrences of {\it zebra} according to its importance as {\sc nsubj}.
% \item {\it Water} as a modifier often changes the meaning of its head in
%  a visually salient way: e.g.\ {\it water fall, water balloon, water scene, water skiing}, and
%   thus the network learns that this particular word is important in the modifier position.
% \end{enumerate}

@@

\begin{figure}
\centering
 \includegraphics[scale=0.4]{position-coef.pdf}
 \caption{Coefficients on the y-axis of {\sc LR full} corresponding to the
position variables on the x-axis.}
 \label{fig:posrqs}
\end{figure}

@@

For the {\sc Visual} model it is less clear what to
expect: on the one hand due to their chain structure,  
RNNs are better at keeping track of
short-distance rather than long-distance dependencies and thus we can expect 
tokens in positions closer to the end of the sentence to be more important. 
On the other hand in English the information structure of a single
sentence is expressed via linear ordering: the {\sc topic} of a 
sentence appears sentence-initially, and the {\sc comment} follows. 
In the context of other text types such as dialog or multi-sentence \label{edit:topiccomment}
narrative structure, we would expect {\sc comment} to often be more
important than {\sc topic} as {\sc comment} will often 
contain new information in these cases. In our setting of image captions 
however, sentences are not part of larger discourse, it is sentence
initial material that typically contains the most important 
objects depicted in the image, e.g. {\it {\underline{two zebras}} are grazing in tall grass on a savannah.}
Thus, for the task of predicting features of the visual scene, it would 
be advantageous to detect the topic of the sentence and up-weight its
importance in the final meaning representation. Figure \ref{fig:posrqs}
appears to support this hypothesis and the network does 
learn to pay more attention to  words appearing
sentence-initially. This effect seems to be to some extent mixed with the recency
bias of RNNs as perhaps indicated by the relatively high coefficient of the {\it last}
position for {\sc Visual}.
